---
title: 'ggplot2: 2. Line Graphs'
author: "suyni00i"
date: "2023-05-13"
output: html_document
---

```{r}
library(ggplot2)
library(gcookbook)
```

# Making a Basic Line Graph
```{r}
str(BOD)
```

```{r}
ggplot(BOD, aes(x = Time, y = demand)) + geom_line()
# x = numeric
```

```{r}
ggplot(BOD, aes(x = factor(Time), y = demand, group = 1)) + geom_line()
# group = 1: every data is grouped as one
```

> geom_line(): line graph

```{r}
ggplot(BOD, aes(x = Time, y = demand)) + geom_line() +
  ylim(0, max(BOD$demand))
# set y-axis range 0~max(BOD%demand)
```

# Adding Points to a Line Graph
```{r}
ggplot(BOD, aes(x = Time, y = demand)) + geom_line() + geom_point()
```
> geom_point(): add points

```{r}
str(worldpop)
```

```{r}
p <- ggplot(worldpop, aes(x = Year, y = Population)) +
  geom_line() + geom_point()
p
```

```{r}
p + scale_y_log10()
# revert y-axis to log10 scale
```

# Making a Line Graph with Multiple Lines
```{r}
str(ToothGrowth)
```

```{r}
library(plyr)
tg <- ddply(ToothGrowth, c("supp", "dose"), summarise, length = mean(len))
tg
```

```{r}
# map supp to color
ggplot(tg, aes(x = dose, y = length, color = supp)) + geom_line()
# color = supp: set color of line accd to supp
```

```{r}
# map supp to linetype
ggplot(tg, aes(x = dose, y = length, color = supp, linetype = supp)) +
  geom_line()
# linetype = supp: set type of line accd to supp
```
> ddply(ToothGrowth, c("supp", "dose"), summarise, length = mean(len)): print dataframe w/ _mean(len)_ accd to comb between _supp_ and _dose_

```{r}
ggplot(tg, aes(x = dose, y = length, shape = supp)) + 
  geom_line() + geom_point(size = 4)
# size: size of the point
```

```{r}
ggplot(tg, aes(x = dose, y = length, fill = supp)) +
  geom_line() + geom_point(size = 4, shape = 21)
# shape: shape of the point, ex. 21 = circle
```
> ggplot(..., shape = supp): differ the shape of the point accd to supp

```{r}
ggplot(tg, aes(x = dose, y = length, shape = supp)) +
  geom_line(position = position_dodge(0.05)) +
  geom_point(position = position_dodge(0.05), size = 4)
```
> position = position_dodge(): array side-by-side, w/o overlap

# Changing the Appearance of Lines and Points
```{r}
ggplot(BOD, aes(x = Time, y = demand)) + 
  geom_line(linetype = "dashed", size = 1, color = "blue") +
  geom_point(size = 10, shape = 22, color = "darkred", fill = "pink")
```

```{r}
ggplot(BOD, aes(x = Time, y = demand)) + geom_line() +
  geom_point(size = 10, shape = 21, fill = "white")
```

# Making a Graph with a Shaded Area
```{r}
str(sunspot.year)
```

```{r}
sunspotyear <- data.frame(
  Year = as.numeric(time(sunspot.year)),
  Sunspots = as.numeric(sunspot.year)
)
ggplot(sunspotyear, aes(x = Year, y = Sunspots)) + geom_area()
```

```{r}
ggplot(sunspotyear, aes(x = Year, y = Sunspots)) +
  geom_area(fill = "blue", alpha = 0.2) + geom_line()
```
> time(sunspot.year): change ts to numeric
> geom_area(): shade area under line
> alpha = 0.2: adjust tansparency(0~1: transparent~opaque)

# Making a Stacked Area Graph
```{r}
str(uspopage)
```

```{r}
ggplot(uspopage, aes(x = Year, y = Thousands, fill = AgeGroup)) + geom_area()
```
> fill = AgeGroup: differ color for each _AgeGroup_

```{r}
p <- ggplot(uspopage, aes(x = Year, y = Thousands, fill = AgeGroup))
p + geom_area(color = "black", size = 0.2, alpha = 0.4) +
  scale_fill_brewer(palette = "Blues", breaks = rev(levels(uspopage$AgeGroup))) + theme_bw()
```
> color = "black": separate areas w/ _black_ lines
> scale_fill_brewer(): set color
> breaks = rev(levels(uspopage$AgeGroup)): reverse _AgeGroup_ level b4 stack
> theme_bw(): set bg color white

```{r}
p <- ggplot(uspopage, aes(x = Year, y = Thousands, fill = AgeGroup))
p + geom_area(color = NA, size = 0.2, alpha = 0.4) +
  scale_fill_brewer(palette = "Blues", breaks = rev(levels(uspopage$AgeGroup))) + theme_bw()
```

```{r}
# removing the border only for legend
p + geom_area(color = NA, size = 0.2, alpha = 0.4) +
  scale_fill_brewer(palette = "Blues", breaks = rev(levels(uspopage$AgeGroup))) +
  geom_line(position = "stack", size = 0.2) + theme_bw()
```
> color = NA: delete boundaries among areas
> geom_line(position = "stack"): add boundaries on graph

# Adding a Confidence Region
```{r}
str(climate)
```

```{r}
clim <- subset(climate, Source == "Berkeley",
               select = c("Year", "Anomaly10y", "Unc10y"))
p <- ggplot(clim, aes(x = Year, y = Anomaly10y))
# shaded region
p + geom_ribbon(aes(ymin = Anomaly10y - Unc10y, ymax = Anomaly10y + Unc10y), alpha = 0.2) + geom_line()
```

```{r}
# dotted lines
p + geom_line(aes(y = Anomaly10y - Unc10y), color = "grey50", linetype = "dotted") +
  geom_line(aes(y = Anomaly10y + Unc10y), color = "grey50", linetype = "dotted") +
  geom_line(lwd = 1)
```
> geom_ribbon(aes(ymin = ___, ymax = ___)): confidence band
