---
title: 'ggplot2: 10. Using Colors in Plots'
author: "suyni00i"
date: "2023-05-17"
output: html_document
---

```{r}
library(ggplot2)
library(gcookbook)
```

# Setting the Colors of Objects
```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) + geom_point(color = "red")
```

```{r}
library(MASS)
ggplot(birthwt, aes(x = bwt)) + geom_histogram(fill = "red", color = "black")
```

# Mapping Variables to Colors
```{r}
# method 1 | "fill" in "aes()"
ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar)) +
  geom_col(color = "black", position = "dodge")
```

```{r}
# method 2 | "aes(fill = )" in "geom"
ggplot(cabbage_exp, aes(x = Date, y = Weight)) +
  geom_col(aes(fill = Cultivar), color = "black", position = "dodge")
```

```{r}
# method 1
ggplot(mtcars, aes(x = wt, y = mpg, color = cyl)) + geom_point()
```

```{r}
# method 2
ggplot(mtcars, aes(x = wt, y = mpg)) + geom_point(aes(color = cyl))
```

# Using a Different Palette for a Discrete Variable
```{r}
str(uspopage)
```

```{r}
p <- ggplot(uspopage, aes(x = Year, y = Thousands, fill = AgeGroup)) + geom_area()
p
```

```{r}
p + scale_fill_discrete()
p + scale_fill_hue()
p + scale_fill_hue(h = c(0, 100))
```
> scale_fill_discrete()/scale_fill_hue(): color of area plot
* scale_fill_hue(h = c(0, 100)): range from _0_ to _100_, devided equally by the number of var. (default: h = c(0, 360))

```{r}
p <- ggplot(uspopage, aes(x = Year, y = Thousands, fill = AgeGroup)) + geom_area()
p + scale_fill_brewer()
# scale_fill_brewer(): (default: "Blues")
```

```{r}
h <- ggplot(heightweight, aes(x = ageYear, y = heightIn, color = sex)) +
  geom_point()
h
```

```{r}
h + scale_color_hue(l = 45)
# scale_color_hue(): set brightness (default: l = 65)
```

```{r}
library(RColorBrewer)
par(mar = c(0, 4, 0, 0))
display.brewer.all()
```

```{r}
p + scale_fill_brewer(palette = "Oranges")
p + scale_fill_brewer(palette = "YlGn")
```

```{r}
p + scale_fill_grey()
p + scale_fill_grey(start = 0.7, end = 0)
```
> scale_fill_grey(): (default: start = 0.2, end = 0.8)
* 0(dark) ≤ start, end ≤ 1(bright)

# Using a Manually Defined Palette for a Discrete Variable
```{r}
h <- ggplot(heightweight, aes(x = ageYear, y = heightIn, color = sex)) +
  geom_point()
h
```

```{r}
h + scale_color_manual(values = c("red", "blue"))
```
> scale_color_manual(): set color name

```{r}
h <- ggplot(heightweight, aes(x = ageYear, y = heightIn, color = sex)) +
  geom_point()
h + scale_color_manual(values = c("#CC6666", "#7777DD"))
```

```{r}
h + scale_color_manual(values = c(m = "darkblue", f = "pink"))
```

```{r}
colors()
# color names on R
```
> scale_color_manual(): set RBG color
> values = c(m = "darkblue", f = "pink"): set color of value on group var.

# Using a Manually Defined Palette for a Continuous Variable
```{r}
p <- ggplot(heightweight, aes(x = ageYear, y = heightIn, color = weightLb)) +
  geom_point(size = 1.5)
p
```

```{r}
p + scale_color_gradient(low = "black", high = "white")
```
> scale_color_gradient(low = "black", high = "white"): set color accd to continuous color var.
* low = _black_, high = _white_ (default: low = "#132B43", high = "#56B1F7")

```{r}
p + scale_color_gradient2(low = "red", mid = "white", high = "blue", midpoint = 110)
```

```{r}
library(scales)
p + scale_color_gradient2(low = muted("red"), mid = "white", high = muted("blue"), midpoint = 110)
```
> scale_color_gradient2(low = "red", mid = "white", high = "blue", midpoint = ): set color accd to continuous color var.
* low = _red_, mid = _white_, high = _blue_ (mid value = _110_)
> muted(colour, 1 = 30, c = 70): show RGB color w/ luminance & chroma adjusted

```{r}
p + scale_color_gradientn(colours = c("darkred", "orange", "yellow", "white"))
p + scale_color_gradientn(colours = c("red", "blue", "green", "black"))
```
> scale_color_gradientn(): set color(4+) accd to continuous color var.