---
title: 'ggplot2: 1. Bar Graphs'
author: "suyni00i"
date: "2023-05-11"
output: html_document
---

```{r}
# library
library(ggplot2)
library(gcookbook)
```

# Making Basic Bar Graph
```{r}
# method 1
ggplot(pg_mean, aes(x = group, y = weight)) +
  geom_col()
```

```{r}
# method 2
p <- ggplot(pg_mean, aes(x = group, y = weight))
p + geom_bar(stat = "identity")
```

```{r}
str(BOD)
ggplot(BOD, aes(x = Time, y = demand)) + geom_col()
```

```{r}
ggplot(BOD, aes(x = factor(Time), y = demand)) + geom_col()
```
> factor: except empty x

```{r}
BOD2 <- BOD
BOD2$fTime <- factor(BOD$Time)
str(BOD)
```

```{r}
ggplot(BOD2, aes(x = fTime, y = demand)) + geom_col()
```

```{r}
ggplot(pg_mean, aes(x = group, y = weight)) +
  geom_col(fill = "lightblue", color = "black")
```

# Grouping Bar Together
```{r}
p <- ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar))
p + geom_col(position = "dodge")
```
> position = "dodge": side-by-side if two
> fill = Cultivar: legend

```{r}
p + geom_col(position = "dodge", color = "black") +
  scale_fill_brewer(palette = "Pastel1")
```
> scale_fill_brewer(): extra option for fill = _"color"_

```{r}
ce <- cabbage_exp
ce[c(1, 6), 3] <-  NA
# regard (1, 3), (6, 3) as missing
ggplot(ce, aes(x = Date, y = Weight, fill = Cultivar)) + 
  geom_col(position = "dodge", color = "black") +
  scale_fill_brewer(palette = "Pastel1")
```

```{r}
ce2 <- cabbage_exp[-c(1, 6), ]
# delete row 1, 6
ggplot(ce2, aes(x = Date, y = Weight, fill = Cultivar)) +
  geom_col(position = "dodge", color = "black") +
  scale_fill_brewer(palette = "Pastel1")
```

# Making a Bar Graph of Counts
```{r}
str(diamonds)
```

```{r}
ggplot(diamonds, aes(x = cut)) + geom_bar()
```

```{r}
ggplot(diamonds, aes(x = carat)) + geom_bar()
```


```{r}
ggplot(diamonds, aes(x = carat)) + geom_histogram()
```
>geom_bar(): w/o y = "___"
* if x = numeric: == geom_histogram()

> geom_histogram()

# Using Colors in a Bar Graph
```{r}
str(uspopchange)
```

```{r}
head(uspopchange)
```

```{r}
upc <- subset(uspopchange, rank(Change) > 40)
upc
```

```{r}
ggplot(upc, aes(x = Abb, y = Change, fill = Region)) + geom_col()
```
> str(data): show R data's structural char.
> head(data): show 6 rows from the front
> tail(data): show 6 rows from the back
> subset(data, condition): select and show _data_ w/ _condition_
> fill = Region: select color accd to _Reigion_

```{r}
ggplot(upc, aes(x = reorder(Abb, Change), y = Change, fill = Region)) +
  geom_col(color = "black") +
  scale_fill_manual(values = c("#669933", "#FFCC66")) +
  xlab("State")
```
> reorder(var1, var2): sort var1 accd to var2
> scale_fill_manual: select colors of fill (#FFGGBB: RGB color code)
> xlab("___"): set x label

# Coloring Negative and Positive Bars Differently
```{r}
csub <- subset(climate, Source == "Berkeley" & Year >= 1900)
# data only w/ Source == "Berkeley" & Year >= 1900
csub$pos <- csub$Anomaly10y >= 0
# pos: TRUE if Anomaly10y >= 0, else FALSE
p <- ggplot(csub, aes(x = Year, y = Anomaly10y, fill = pos))
p + geom_col(position = "identity")
```

```{r}
p <- ggplot(csub, aes(x = Year, y = Anomaly10y, fill = pos))
p + geom_col(position = "identity", color = "black", size = 0.25) +
  scale_fill_manual(values = c("#CCEEFF", "#FFDDDD"), guide = FALSE)
```
> size = ___: set bar's width
> guide = FALSE: no legend printed

# Adjusting Bar Width and Spacing
```{r}
p <- ggplot(pg_mean, aes(x = group, y = weight))
p + geom_col() # default width = 0.9
```

```{r}
p + geom_col(width = 0.5) # narrower
p + geom_col(width = 1) # wider
```

```{r}
p <- ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar))
p + geom_col(width = 0.5, position = "dodge")
```

```{r}
p + geom_col(width = 0.5, position = position_dodge(0.7))
```
> position = "dodge" == position = position_dodge( _value of width_ )

# Making Stacked Bar Graph
```{r}
p <- ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar))
p + geom_col()
```

```{r}
p + geom_col() + guides(fill = guide_legend(reverse = TRUE))
# reverse legend
```

# Adding Labels to a Bar Graph
```{r}
p <- ggplot(cabbage_exp, aes(x = interaction(Date, Cultivar), y = Weight)) +
  geom_col()
p + geom_text(aes(label = Weight))
```

```{r}
p + geom_text(aes(label = Weight), vjust = 1.5, color = "white")
```

```{r}
p + geom_text(aes(label = Weight), vjust = -0.2)
```
> interaction(Date, Cultivate): interaction btw _Date, Cultivate_
> geom_text(): print label
* vjust: move the loc of text
