---
title: 'ggplot2: 3. Scatter Plots'
author: "suyni00i"
date: "2023-05-14"
output: html_document
---

```{r}
library(ggplot2)
library(gcookbook)
```

# Making a Basic Scatter Plots
```{r}
str(heightweight)
```

```{r}
head(heightweight[, c("ageYear", "heightIn")])
```

```{r}
p <- ggplot(heightweight, aes(x = ageYear, y = heightIn))
p + geom_point(shape = 21)
```

```{r}
p + geom_point(sizq = 1.5)
```
# Grouping Data Points by a Variable Using Shape or Color
```{r}
head(heightweight[, c("sex", "ageYear", "heightIn")])
```

```{r}
p <- ggplot(heightweight, aes(x = ageYear, y = heightIn, shape = sex, color = sex))
p + geom_point()
```

```{r}
p + geom_point() + scale_shape_manual(values = c(1, 2)) +
  scale_color_brewer(palette = "Set1")
```

# Using Different Point Shapes
```{r}
ggplot(heightweight, aes(x = ageYear, y = heightIn)) + geom_point(shape = 3)
```

```{r}
ggplot(heightweight, aes(x = ageYear, y = heightIn, shape = sex)) +
  geom_point(size = 4) + scale_shape_manual(values = c(1, 4))
```

```{r}
hw <- heightweight
hw$weightGroup <- cut(hw$weightLb, breaks = c(-Inf, 100, Inf),
                      labels = c("< 100", ">= 100"))
ggplot(hw, aes(x = ageYear, y = heightIn, shape = sex, fill = weightGroup)) +
  geom_point(size = 2.5) +
  scale_shape_manual(values = c(21, 24)) +
  scale_fill_manual(values = c(alpha("black", 0), "black"),
                    guide = guide_legend(override.aes = list(shape = 21)))
```
> cut(): make weightGroup
> scale_shape_manual(): shape of point
> scale_fill_manual(): color of point

# Mapping a Continuous Variable to Color or Size
```{r}
ggplot(heightweight, aes(ageYear, y = heightIn, color = weightLb)) +
  geom_point()
```

```{r}
ggplot(heightweight, aes(x = ageYear, y = heightIn, size = weightLb)) +
  geom_point()
```

```{r}
ggplot(heightweight, aes(x = ageYear, y = heightIn, size = weightLb)) +
  geom_point()
```

```{r}
ggplot(heightweight, aes(x = ageYear, y = heightIn, fill = weightLb)) +
  geom_point(shape = 21, size = 2.5) +
  scale_fill_gradient(low = "black", high = "white")
```

```{r}
ggplot(heightweight, aes(x = ageYear, y = heightIn, fill = weightLb)) +
  geom_point(shape = 21, size = 2.5) +
  scale_fill_gradient(low = "black", high = "white", guide = guide_legend())
```
> scale_fill_gradient(low = "black", high = "white"): _black_ if low, _white_ if high
> guide = guide_legend(): discrete legend

```{r}
ggplot(heightweight, aes(x = ageYear, y = heightIn, size = weightLb, color = sex)) +
  geom_point(alpha = 0.5) +
  scale_size(range = c(1, 6)) +
  scale_color_brewer(palette = "Set1")
```

# Adding Fitted Regression Model Lines
```{r}
sp <- ggplot(heightweight, aes(x = ageYear, y = heightIn))
sp + geom_point() + stat_smooth(method = lm)
```

```{r}
sp + geom_point() + stat_smooth(method = lm, level = 0.99)
```

```{r}
sp + geom_point() + stat_smooth(method = lm, se = FALSE)
```

```{r}
sp + geom_point() + stat_smooth()
# default: method = loess
```
> stat_smooth(): scatter plot + regression(method = lm)
> level = 0.99: 99% confidence interval
> se = FALSE: w/o confidence interval

```{r}
sps <- ggplot(heightweight, aes(x = ageYear, y = heightIn, color = sex)) +
  geom_point() + scale_color_brewer(palette = "Set1")
sps + stat_smooth(method = lm)
```

```{r}
sps + stat_smooth(method = lm, fullrange = TRUE)
# fullrange = TRUE: set full range as widest case
```

# Adding Annotations with Model Coefficients
```{r}
model <- lm(heightIn ~ ageYear, data = heightweight)
summary(model)
```

```{r}
sp <- ggplot(heightweight, aes(x = ageYear, y = heightIn)) + geom_point() +
  stat_smooth(method = lm)
sp + annotate("text", label = "r^2 == 0.42", parse = TRUE, x = 16.5, y = 52)
```

```{r}
est <- unname(coef(model))
eqn <- as.character(as.expression(
  substitute(italic(y) == a + b * italic(x) * "," ~~ italic(r)^2 ~ "=" ~ r2,
             list(a = format(est[1], digits = 3),
                  b = format(est[2], digits = 3),
                  r2 = format(summary(model)$r.squared, digits = 2)
))))
parse(text = eqn)
```

```{r}
sp + annotate("text", label = eqn, parse = TRUE, x = Inf, y = -Inf, hjust = 1.1, vjust = -0.5)
```
> annotate(): show text, formulas, shapes, ... on graph
* c.f geom(): mapping

# Adding Marginal Rugs to a Scatter Plot
```{r}
ggplot(faithful, aes(x = eruptions, y = waiting)) + geom_point() + geom_rug()
```

```{r}
ggplot(faithful, aes(x = eruptions, y = waiting)) + geom_point() + geom_rug(position = "jitter", size = 0.2)
```
> geom_rug(): add rug plot on each axis
> position = "jitter": add random noise on overlapped points -> show density

# Labeling Points in a Scatter Plots
```{r}
cs <- subset(countries, Year == 2009 & healthexp > 2000)
str(cs)
```

```{r}
sp <- ggplot(cs, aes(x = healthexp, y = infmortality)) + geom_point()
sp + geom_text(aes(label = Name), size = 4)
```

```{r}
sp + geom_text(aes(label = Name), size = 4, vjust = -1)
```

```{r}
sp + geom_text(aes(label = Name), size = 4, hjust = -0.1)
```
> geom_text(label = Name): show _Name_ as text label
> vjust, hjust: edit loc of text label

```{r}
sp + annotate("text", x = 4350, y = 5.4, label = "Canada") +
  annotate("text", x = 7400, y = 6.8, label = "USA")
```

```{r}
cs$Name1 <- cs$Name
idx <- cs$Name1 %in% c("Canada", "Ireland", "United Kingdom", "United States", "Japan", "New Zealand", "Iceland", "Luxembourg", "Netherland", "Switzerland")
cs$Name1[!idx] <- NA
ggplot(cs, aes(x = healthexp, y = infmortality, label = Name1)) + geom_point() +
  geom_text(aes(x = healthexp + 100, lable = Name1), size = 4, hjust = 0) +
              xlim(2000, 9000)
```
> vec1 %in% vec2: select same elements as _vec2_ among _vec1_
* (is.element(vec1, vec2))

# Creating a Balloon Plot
```{r}
cdat <- subset(countries, Year == 2009 &
                 Name %in% c("Canada", "Ireland", "United Kingdom", "United States", "New Zealand", "Iceland", "Japan", "Luxembourg", "Netherland", "Switzerland"))
p <- ggplot(cdat, aes(x = healthexp, y = infmortality, size = GDP)) +
  geom_point(shape = 21, color = "black", fill = "cornsilk")
p
```

```{r}
p + scale_size_area(max_size = 15)
```

```{r}
p + scale_size(range = c(3, 10))
# range = c(3, 10): set biggest point's size
```