---
title: 'ggplot2: 6. Axes'
author: "suyni00i"
date: "2023-05-16"
output: html_document
---

```{r}
library(ggplot2)
library(gcookbook)
```

# Swapping X- and Y-Axes
```{r}
ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot()
```

```{r}
ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot() + coord_flip()
```
> coord_flip(): Change two Axes(x <-> y)
* variation = before change

```{r}
ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot() + coord_flip() +
  scale_x_discrete(limits = rev(levels(PlantGrowth$group)))
```
> scale_x_discrete(): if discrete(factor) var., graphic option on x-axis
* if coord_flip(): " on y-axis

# Setting the Range of a Continuous Axis
```{r}
p <- ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot()
p
```

```{r}
p + ylim(0, max(PlantGrowth$weight))
```

```{r}
# ylim(): shorthand for setting the limits w/ scale_y_continuous()
p + scale_y_continuous(limits = c(0, max(PlantGrowth$weight)))
```
> scale_y_continuous(): if continuous(numeric) var., graphic option on y-axis
* ylim(): if only range of axis

```{r}
p <- ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot()
p + scale_y_continuous(limits = c(5, 6.5))
# redraw boxplots within the limit
# use only data matches
```

```{r}
p + coord_cartesian(ylim = c(5, 6.5))
# magnify the limit region
# use total data, show limited areas on sheet(magnified)
```

# Reversing a Continuous Axis
```{r}
p <- ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot()
p + scale_y_reverse()
```
> scale_y_reverse(): show data on y-axis reversed

# Changing the Order of Items on a Categorical Axis
```{r}
g <- ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot()
p + scale_x_discrete(limits = c("trt1", "ctrl", "trt2"))
```

```{r}
p + scale_x_discrete(limits = c("ctrl", "trt1"))
```
> scale_x_discrete(limits = ): change order of var./show limited area
* categorical data

# Setting the Positions of Tick Marks
```{r}
p <- ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot()
p + scale_y_continuous(breaks = c(4, 4.25, 4.5, 5, 6, 8))
```

```{r}
mytick <- seq(3.5, 7, by = 0.2)
p + scale_y_continuous(breaks = mytick) +
  scale_x_discrete(limits = c("ctrl", "trt1"), breaks = "crtl")
```
> scale_x_discrete(): set tick marks of categorical var.
> scale_y_continuous(): set tick marks of continuous var.

# Removing Tick Marks and Labels
```{r}
p <- ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot()
p + theme(axis.text.y = element_blank())
# don't show tick label of y-axis
```

```{r}
p + theme(axis.ticks = element_blank())
# don't show tick label of every axes
```

```{r}
p <- ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot()
p + theme(axis.ticks = element_blank(), axis.text.y = element_blank())
# remove tick label of every axes
```

```{r}
p + scale_y_continuous(breaks = NULL)
# remove grids on y
```

# Changing the Text of Tick Labels
```{r}
hwp = ggplot(heightweight, aes(x = ageYear, y = heightIn)) + geom_point()
hwp
```

```{r}
hwp + scale_y_continuous(breaks = c(50, 56, 60, 66, 72),
                         labels = c("Tiny", "Really\nshort", "Short", "Medium", "Tallish"))
# change labels
```

```{r}
footinch_formatter <- function (x) {
  foot <- floor(x / 12)
  inch <- x %% 12
  return( paste(foot, "'", inch, "\"", sep = ""))
}
footinch_formatter(56:64)
# function for inch -> feet/inch
# %%: remainder of division
```

```{r}
hwp + scale_y_continuous(labels = footinch_formatter)
```

```{r}
hwp + scale_y_continuous(breaks = seq(48, 72, by = 4), labels = footinch_formatter)
```
> floot(): int ≤ function factor

```{r}
timeHMS_formatter <- function (x) {
  h <- floor(x / 60) # hour
  m <- floor(x %% 60) # minute
  s <- round(60 * (x %% 1)) # Round to nearest second
  lab <- sprintf("%02d:%02d:%02d", h, m, s) # format the strings as HH:MM:SS
  lab <- gsub("^00:", "", lab) # Remove leading 00: if present
  lab <- gsub("^0", "", lab) # Remove leading 0 if present
  return (lab)
}
timeHMS_formatter(c(.33, 50, 51, 51.25, 60, 60.1, 130.23, 1234))
# timeHMS_formatter: return time as HH/MM/SS
```

# Changing the Appearance of Tick Labels
```{r}
bp <- ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot() +
  scale_x_discrete(breaks = c("ctrl", "trt1", "trt2"),  labels = c("Control", "Treatment 1", "Treatment 2"))
bp
```

```{r}
bp + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

```{r}
bp + theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))
```

```{r}
bp + theme(axis.text.x = element_text(family = "Times", face = "italic", color = "darkred", size = rel(2)))
```

```{r}
library(extrafont)
```

```{r}
# font_import()
# ... fail
```

```{r}
# fonts()
```
> axis.text.x = element_text(): change shape of tick label
* angle, hjust&vjust(loc.), family(font), face(typeset)

# Changing the Text of Axis Labels
```{r}
hwp <- ggplot(heightweight, aes(x = ageYear, y = heightIn, color = sex)) +
  geom_point()
hwp
```

```{r}
# set the axis labels
hwp + xlab("Age in Years") + ylab("Height in inches")
```

```{r}
hwp + labs(x = "Age in years", y = "Height in inches")
```

```{r}
hwp + scale_y_continuous(name = "Height (inches)") + scale_x_continuous(name = "Age\n(years)")
```
> xlab(), ylab(), labs(), scale_x_continuous(), scale_y_continuous(): change labels

# Removing Axis Labels
```{r}
p <- ggplot(PlantGrowth, aes(x = group, y = weight)) + geom_boxplot()
p
```

```{r}
p + theme(axis.title.x = element_blank())
```

```{r}
p + xlab("")
```
> xlab(""), theme(axis.title.x = element_blank()): remove label of x-axis

# Changing the Appearance of Axis Labels
```{r}
hwp <- ggplot(heightweight, aes(x = ageYear, y = heightIn, color = sex)) +
  geom_point()
hwp + theme(axis.title.x = element_text(face = "italic", color = "darkred", size = 14)) +
  theme(axis.title.y = element_text(face = "italic", angle = 90, color = "darkred", size = 14)) +
  ylab("Height\n(Inches)")
```
> element_text(): Edit axis label

# Showing Lines Along the Axes
```{r}
p <- ggplot(heightweight, aes(x = ageYear, y = heightIn)) + geom_point()
p
```

```{r}
p + theme(axis.line = element_line(color = "black"))
p + theme_bw()
```

```{r}
p + theme_bw() + theme(panel.border = element_blank(), panel.grid = element_blank(),
                       axis.line = element_line(color = "black"))
```

```{r}
p + theme_bw() + theme(panel.border = element_blank(),
                       axis.line = element_line(color = "black", size = 4, lineend = "square"))
```