---
title: 'ggplot2: 9. Facets'
author: "suyni00i"
date: "2023-05-17"
output: html_document
---

```{r}
library(ggplot2)
library(gcookbook)
```

# Splitting Data into Subplots with Facets
```{r}
str(mpg)
```

```{r}
p <- ggplot(mpg, aes(x = displ, y = hwy)) + geom_point(size = 0.8)
# faceted by drv, in vertically arranged subpanels
p + facet_grid(drv ~ .)
```

```{r}
# faceted byb cyl, in horizontally arranged subpanels
p + facet_grid(. ~ cyl)
```
> facet_grid(drv ~ .): set facet vertical accd to drv
> facet_grid(. ~ cyl): set facet horizontal accd to cyl

```{r}
p <- ggplot(mpg, aes(x = displ, y = hwy)) + geom_point(size = 0.8)
p + facet_grid(drv ~ cyl)
```
> facet_grid(drv ~ cyl): scatterplot btw _displ_ & _hwy_
* horizontal axis = _drv_, vertical axis = _cyl_

```{r}
p <- ggplot(mpg, aes(x = displ, y = hwy)) + geom_point(size = 0.8)
p + facet_wrap(~ class, nrow = 2)
```

```{r}
p + facet_wrap(~ class, ncol = 5)

```
> facet_wrap(~ class, nrow = 2): scatterplot btw _displ_ & _hwy_
* array: row = _2_
> facet_wrap(~ class, ncol = 5): scatterplot btw _displ_ & _hwy_
* array: column = _5_

# Using Fact with Different Axes
```{r}
p <- ggplot(mpg, aes(x = displ, y = hwy)) + geom_point(size = 0.8)
p + facet_grid(drv ~ cyl, scales = "free_y")
# scales = "free_y": re-scale vertical-axis of facets
```

```{r}
p + facet_grid(drv ~ cyl, scales = "free")
# scales = "free":  re-scale vertical & horizontal axes of facets
```

# Changing the Text of Facet Labels
```{r}
mpg2 <- mpg
levels(mpg2$drv)
```

```{r}
levels(mpg2$drv) <- c("4wd", "Front", "Rear")
ggplot(mpg2, aes(x = displ, y = hwy)) + geom_point(size = 0.5) + facet_grid(drv ~ .)
```

# Changing the Appearance of Facet Labels
```{r}
ggplot(cabbage_exp, aes(x = Cultivar, y = Weight)) +
  geom_col() +
  facet_grid(. ~ Date) +
  theme(strip.text = element_text(face = "bold", size = rel(1.5), color = "red"),
        strip.background = element_rect(fill = "lightblue", color = "black", size = 1))
```